---
title: "Dunhumby Explore_User Responde Model"
output: html_document
date: "2025-04-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load the dataset
```{r}
# Load libraries
library(tidyverse)

# Set data path
data_path <- "/Users/shimuliu/Desktop/First-year Paper/archive"

# Load datasets
campaign_desc       <- read_csv(file.path(data_path, "campaign_desc.csv"))
campaign_table      <- read_csv(file.path(data_path, "campaign_table.csv"))
causal_data         <- read_csv(file.path(data_path, "causal_data.csv"))
coupon_redempt      <- read_csv(file.path(data_path, "coupon_redempt.csv"))
coupon              <- read_csv(file.path(data_path, "coupon.csv"))
hh_demographic      <- read_csv(file.path(data_path, "hh_demographic.csv"))
product             <- read_csv(file.path(data_path, "product.csv"))
transaction_data    <- read_csv(file.path(data_path, "transaction_data.csv"))
```
2.Join transaction and product tables
```{r}
# Join transaction with product information
transaction_with_product <- transaction_data %>%
  left_join(product, by = "PRODUCT_ID")

#check unknowns
sum(is.na(transaction_with_product$DEPARTMENT))
colSums(is.na(transaction_with_product))

#Replace unkown department with NA
transaction_with_product <- transaction_with_product %>%
  mutate(DEPARTMENT = ifelse(is.na(DEPARTMENT), "Unknown", DEPARTMENT))

#Replace unknown COMMODITY_DESC with NA
transaction_with_product <- transaction_with_product %>%
  mutate(COMMODITY_DESC = ifelse(is.na(COMMODITY_DESC), "Unknown", COMMODITY_DESC))

#Replace SUB_COMMODITY_DESC with NA
transaction_with_product <- transaction_with_product %>%
  mutate(SUB_COMMODITY_DESC = ifelse(is.na(SUB_COMMODITY_DESC), "Unknown", SUB_COMMODITY_DESC))

```
7839 missing values in DEPARTMENT, COMMODITY_DESC and SUB_COMMODITY_DESC. 

3. Compute total sales of each department (sales in dollars)
```{r}
#Create a new dataset to store department sales
department_sales <- transaction_with_product %>%
  group_by(DEPARTMENT) %>%
  summarise(TOTAL_DEPARTMENT_SALES = sum(SALES_VALUE, na.rm = TRUE)) %>%
  arrange(desc(TOTAL_DEPARTMENT_SALES))

# Summarize the department names only
department_summary <- transaction_with_product %>%
  count(DEPARTMENT, sort = TRUE)

# View the result
print(department_summary)

#Inspect
summary(department_sales)
head(department_sales, 10)
```

4. Pick a category with high sales
```{r}
# Select a department with high sales
top_10_departments <- department_sales %>%
  arrange(desc(TOTAL_DEPARTMENT_SALES)) %>%
  slice_head(n = 10)

# View result
print(top_10_departments)
```
Grocery department has the highest sales, followed by Produce and Frozen Foods.

5. Find the brand with th highest sales
```{r}

# 1. Filter for only Grocery department
grocery_data <- transaction_with_product %>%
  filter(DEPARTMENT == "GROCERY")

# 2. Group by BRAND and summarize total sales
brand_sales_in_grocery <- grocery_data %>%
  group_by(BRAND) %>%
  summarise(TOTAL_BRAND_SALES = sum(SALES_VALUE, na.rm = TRUE)) %>%
  arrange(desc(TOTAL_BRAND_SALES))

# 3. Pick the top brand
top_brand_in_grocery <- brand_sales_in_grocery %>%
  slice_head(n = 5)

# View the result
print(top_brand_in_grocery)

```


6. Find the SUB_COMMODITY_DESC with the highest sales in the Grocery department
```{r}
# 1. Filter for only Grocery department
grocery_data <- transaction_with_product %>%
  filter(DEPARTMENT == "GROCERY")

# 2. Group by BRAND and summarize total sales
commodity_desc_sales_in_grocery <- grocery_data %>%
  group_by(SUB_COMMODITY_DESC) %>%
  summarise(SUB_COMMODITY_DESC_SALES = sum(SALES_VALUE, na.rm = TRUE)) %>%
  arrange(desc(SUB_COMMODITY_DESC_SALES))

# 3. Pick the top brand
top_commodity_in_grocery <- commodity_desc_sales_in_grocery %>%
  slice_head(n = 5)

# View the result
print(commodity_desc_sales_in_grocery)
```
The top two sub-commodities in the grocery department is FLUID MILK WHITE ONLY and SOFT DRINKS 12/18&15PK CAN CAR. 

7. Check if the top sub-commodities have ads presented
```{r}
# 1. Get all PRODUCT_IDs belonging to that SUB_COMMODITY_DESC
top_commodity_products <- transaction_with_product %>%
  filter(DEPARTMENT == "GROCERY", SUB_COMMODITY_DESC == "FLUID MILK WHITE ONLY") %>% 
  distinct(PRODUCT_ID)

# 2.From causal_data, filter only those products
top_commodity_causal <- causal_data %>%
  filter(PRODUCT_ID %in% top_commodity_products$PRODUCT_ID)

# 3. Summarize number of different displays
display_summary <- top_commodity_causal %>%
  count(display, sort = TRUE)

print(display_summary)

# 4. Summarize number of different mailers
mailer_summary <- top_commodity_causal %>%
  count(mailer, sort = TRUE)

print(mailer_summary)
```
8. Prepare the data for user-level sales response model
Y_{ist} = sales of "FLUID MILK WHITE ONLY" to household i in store s in week t
X_{ist} = price to household i in store s in week t  - tricky what price did they face when they don’t buy?
        display
        mailer
        coupon
      
8.a.Set target sub-commodity "FLUID MILK WHITE ONLY"
```{r}
target_sub_commodity <- "FLUID MILK WHITE ONLY"

# Get all Product IDs under this sub-commodity
target_products <- product %>%
  filter(SUB_COMMODITY_DESC == target_sub_commodity) %>%
  distinct(PRODUCT_ID)

print(target_products)
```

8.b.Build the full dataset: Sales of Household * Store * Week
```{r}
# Get all unique households, stores, and weeks
households <- transaction_data %>% distinct(household_key)
stores <- transaction_data %>% distinct(STORE_ID)
weeks <- transaction_data %>% distinct(WEEK_NO)

# Expand grid: Household x Store x Week
panel <- crossing(households, stores, weeks)

print(panel)
```
8.c. Merge sales data
```{r}
# Filter transactions for target milk products
milk_sales <- transaction_data %>%
  filter(PRODUCT_ID %in% target_products$PRODUCT_ID) %>%
  group_by(household_key, STORE_ID, WEEK_NO) %>%
  summarise(
    SALES_VALUE = sum(SALES_VALUE, na.rm = TRUE),
    QUANTITY = sum(QUANTITY, na.rm = TRUE),
    RETAIL_DISC = sum(RETAIL_DISC, na.rm = TRUE),
    COUPON_DISC = sum(COUPON_DISC, na.rm = TRUE),
    COUPON_MATCH_DISC = sum(COUPON_MATCH_DISC, na.rm = TRUE),
    .groups = "drop"
  )

print(milk_sales)
```
8.d. Merge with the full panel
```{r, eval=FALSE}
panel <- panel %>%
  left_join(milk_sales, by = c("household_key", "STORE_ID", "WEEK_NO")) %>%
  mutate(
    SALES_VALUE = replace_na(SALES_VALUE, 0),
    QUANTITY = replace_na(QUANTITY, 0),
    RETAIL_DISC = replace_na(RETAIL_DISC, 0),
    COUPON_DISC = replace_na(COUPON_DISC, 0),
    COUPON_MATCH_DISC = replace_na(COUPON_MATCH_DISC, 0)
  )

print(panel)

```
smaller left join
```{r, eval=FALSE}
# Loop over weeks
for (w in weeks) {
  cat("Processing week", w, "\n")
  
  panel_small <- panel %>% filter(WEEK_NO == w)
  
  joined_small <- panel_small %>%
    left_join(milk_sales, by = c("household_key", "STORE_ID", "WEEK_NO"))
  
  # Check and add missing columns if necessary
  if (!"SALES_VALUE" %in% names(joined_small)) joined_small$SALES_VALUE <- NA
  if (!"QUANTITY" %in% names(joined_small)) joined_small$QUANTITY <- NA
  if (!"RETAIL_DISC" %in% names(joined_small)) joined_small$RETAIL_DISC <- NA
  if (!"COUPON_DISC" %in% names(joined_small)) joined_small$COUPON_DISC <- NA
  if (!"COUPON_MATCH_DISC" %in% names(joined_small)) joined_small$COUPON_MATCH_DISC <- NA
  
  # Now safe to replace NA with 0
  joined_small <- joined_small %>%
    mutate(
      SALES_VALUE = replace_na(SALES_VALUE, 0),
      QUANTITY = replace_na(QUANTITY, 0),
      RETAIL_DISC = replace_na(RETAIL_DISC, 0),
      COUPON_DISC = replace_na(COUPON_DISC, 0),
      COUPON_MATCH_DISC = replace_na(COUPON_MATCH_DISC, 0)
    )
  
  result_list[[as.character(w)]] <- joined_small
}

# After all weeks processed, bind them together
panel_final <- bind_rows(result_list)
head(panel_final)
```
sampling a subset only 500 households
```{r}
# Set seed 
set.seed(123)

# Sample 500 households (you can change this number if needed)
households_small <- households %>% sample_n(500)

# Expand grid: Household x Store x Week (only sampled households)
panel_small <- crossing(households_small, stores, weeks)

# Check size
print(dim(panel_small))
```
```{r}
# Merge the small panel with milk sales
panel_small_joined <- panel_small %>%
  left_join(milk_sales, by = c("household_key", "STORE_ID", "WEEK_NO"))

# Replace NAs with 0s
needed_cols <- c("SALES_VALUE", "QUANTITY", "RETAIL_DISC", "COUPON_DISC", "COUPON_MATCH_DISC") 
panel_small_joined <- panel_small_joined %>%
  mutate(across(all_of(needed_cols), ~replace_na(., 0)))

# Check
summary(panel_small_joined)

```

8.e. Merge with price data (replace NAs with store average price at that week if a sale has happened in that store at that week, then NAs with average store price during all periods if no sales happened during that week but at other weeks, and drop stores with no sales at all bcs it means that store does not sell milk)
```{r}
# Compute average price at each store-week (across all milk buyers)
avg_price_store_week <- transaction_data %>%
  filter(PRODUCT_ID %in% target_products$PRODUCT_ID, QUANTITY > 0) %>%
  mutate(unit_price = SALES_VALUE / QUANTITY) %>%
  group_by(STORE_ID, WEEK_NO) %>%
  summarise(avg_price = mean(unit_price, na.rm = TRUE), .groups = "drop")

# Preview
print(head(avg_price_store_week))
```
```{r}
# Compute overall store-level average price 
avg_price_store <- transaction_data %>%
  filter(PRODUCT_ID %in% target_products$PRODUCT_ID, QUANTITY > 0) %>%
  mutate(unit_price = SALES_VALUE / QUANTITY) %>%
  group_by(STORE_ID) %>%
  summarise(store_avg_price = mean(unit_price, na.rm = TRUE), .groups = "drop")

print(head(avg_price_store))
```
```{r}
# Merge store-week avg price into panel 
panel_small_joined <- panel_small_joined %>%
  left_join(avg_price_store_week, by = c("STORE_ID", "WEEK_NO"))

# Merge overall store avg price into panel 
panel_small_joined <- panel_small_joined %>%
  left_join(avg_price_store, by = "STORE_ID")

# Fill missing avg price: first store-week, then store avg 
panel_small_joined <- panel_small_joined %>%
  mutate(
    avg_price_filled = coalesce(avg_price, store_avg_price)
  )

# Drop stores that have no milk sales ever
panel_small_joined <- panel_small_joined %>%
  filter(!is.na(avg_price_filled))

# Calculate final price variable 
panel_small_joined <- panel_small_joined %>%
  mutate(
    observed_price = if_else(QUANTITY > 0, SALES_VALUE / QUANTITY, NA_real_),
    price = coalesce(observed_price, avg_price_filled)
  )

#Check 
summary(panel_small_joined$price)
sum(is.na(panel_small_joined$price))
```

9. Merge ad types data
```{r}
#From causal_data, filter only our targeted products
causal_milk <- causal_data %>%
  filter(PRODUCT_ID %in% target_products$PRODUCT_ID)

# Collapse to store-week level: if any product is displayed or mailed, mark it as 1
causal_milk_store_week <- causal_milk %>%
  group_by(STORE_ID, WEEK_NO) %>%
  summarise(
    display = max(display, na.rm = TRUE),
    mailer = max(mailer, na.rm = TRUE),
    .groups = "drop"
  )

# Preview
print(causal_milk_store_week)
```
```{r}
#Merge into main
panel_small_joined <- panel_small_joined %>%
  left_join(causal_milk_store_week, by = c("STORE_ID", "WEEK_NO"))

#Replace NAs with 0 (is this reasonable?)
panel_small_joined <- panel_small_joined %>%
  mutate(
    display = if_else(is.na(display), "0", as.character(display)),
    mailer = if_else(is.na(mailer), "0", as.character(mailer))
  )

#Check
panel_small_joined %>%
  filter(STORE_ID == 286, WEEK_NO %in% c(9, 11, 13, 14,21)) %>%
  select(display, mailer)
```

10. User response Model
Yist = SALES_VALUE ist
Xist = price ist
        display ist
        RETAIL_DISC ist
        COUPON_DISC ist
        COUPON_MATCH_DISC ist
```{r}
#Convert to factors
panel_small_joined <- panel_small_joined %>%
  mutate(
    display_cat = factor(display),
    mailer_cat = factor(mailer)
  )

# Fit the model
ols_model_cat <- lm(
  SALES_VALUE ~ price + display_cat + mailer_cat +
    RETAIL_DISC + COUPON_DISC + COUPON_MATCH_DISC,
  data = panel_small_joined
)

options(scipen = 999)
summary(ols_model_cat)

```

